Equalizer: A Mature Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
References and related work
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@19 db7c848b-7755-4df2-b763-8796797afa52

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention. While some commercial
implementations such as TechViz and MechDyne Conduit continue to exists, only
ClusterGL?~\cite{NHM:11} has been published. ClusterGL employs the same approach
as Chromium, but delivers a significantly faster implementation of transparent
OpenGL interception and distribution for parallel rendering.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention. Sage~\cite{TBD} and Sage2~\cite{TBD}

@@

Various applications and frameworks have used Equalizer for new research in
visualization, such as the RTNeuron~\cite{HBBES:13}, an OpenGL renderer for
visualizing simulations of neucortical brain regions, Omegalib~\cite{TBD},
extending Equalizer for the hybrid 2D/3D environments such as the CAVE2,
Siegen~\cite{TBD}

@@

\cite{TBD} DisplayCluster

Omegalib

@@

Tiled interactive display environment, parallel pixel streaming, events

@@

\subsection{Livre}
\subsection{RTT Deltagen}

>>>>

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention since 2009. While some
commercial implementations such as TechViz and MechDyne Conduit continue to
exist, on the research side only ClusterGL~\cite{NHM:11} has been
presented. ClusterGL employs the same approach as Chromium, but delivers a
significantly faster implementation of transparent OpenGL interception and
distribution for parallel rendering. CGLX~\cite{CGLX:11} tries to bring parallel
execution transparently to OpenGL applications, by emulating the GLUT API and
intercepting certain OpenGL calls. This approach works transparently for trivial
applications, but quickly requires the application developer to address the
complexities of a distributed application when mutable application state needs
to be synchronized across processes.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention. Sage~\cite{Sage} and Sage2~\cite{Sage2} have
received significant attention. Sage is build entirely around the concept of a
shared framebuffer where all content windows are separate applications using
pixel streaming. It is no longer actively supported. Sage 2 is a complete,
browser-centric reimplementation where each application is a web application
distributed across browser instances. DisplayCluster~\cite{DC}, and its
continuation Tide, also implement the shared framebuffer concept of Sage, but
provide a few native content applications integrated into the display servers.

@@

Various applications and frameworks have used Equalizer for new research in
visualization, such as the RTNeuron~\cite{HBBES:13}, an OpenGL renderer for
visualizing simulations of neucortical brain regions, Omegalib~\cite{Omegalib},
extending Equalizer for the hybrid 2D/3D environments such as the CAVE2,
visualization for remote sensing date~\cite{LK:09}, .

@@

@@

Tide (Tiled interactive display environment) is an improved version of
DisplayCluster~\cite{DC}, providing touch-based, multi-window user interface for
high-resolution tiled display walls. Remote applications receive input events
and send pixel streams using the Deflect client library. Equalizer includes full
support, enabling application-transparent integration with Tide. When a Tide
server is configured, all output channels of a view stream in parallel to one
window on the wall. In \cite{deflect} we have shown interactive framerates for
an 24 megapixel resolution over a WAN link. Events received are translated into
Equalizer events and injected into the normal event flow, allowing seamless
application integration.

@@

In this section, we present some major applications built using Equalizer, and
show how they interact with the framework to solve complex parallel rendering
problems.

\subsection{Livre}

Livre (Large-scale interactive volume rendering engine) is an out-of-core
parallel 4D volume renderer.

\subsection{RTT Deltagen}
